Analysis and results

### 2020-09-10 analysis/03_analysis.Rmd
Model output

```{r load-packages, warning=FALSE, message=FALSE}
library(tidyverse)
library(donorheuristics)
library(rstan)
library(bayesplot)
library(tidybayes)
library(ggridges)
library(ggraph)
library(ggdag)
library(here)

@@

After selecting the model we'd like to run, model calibration proceeds as follows. Note that model calibration is run in a separate R script given how long it takes to run each model. Model output can be loaded for analysis in the next section.

@@

The model is calibrated using a non-centered parameterization of a hierarchical multinomial logit model.

@@

# Load data and model output

Once we specify the model we'd like to evaluate, the following analysis follows.

```{r load-data-output, warning=FALSE, message=FALSE}
intercept <- 0                            # Intercept-only
public_affairs <- 0                       # Public affairs
political_ideology <- 0                   # Political ideology
social_views <- 0                         # Social views
charity_voluntarism <- 0                  # Charity and voluntarism
demographics <- 0                         # Demographics
public_political <- 0                     # Public affairs + Political ideology
public_political_social <- 0              # + Social views
public_political_social_charity <- 0      # + Charity and voluntarism
public_political_social_charity_demo <- 1 # + Demographics

# Load and extract data and model output (see model-calibration.R)
if (intercept == 1) run <- read_rds(here::here("analysis", "output", "model_runs", "intercept_noncentered.rds"))
if (public_affairs == 1) run <- read_rds(here::here("analysis", "output", "model_runs", "public_affairs.rds"))
if (political_ideology == 1) run <- read_rds(here::here("analysis", "output", "model_runs", "political_ideology.rds"))
if (social_views == 1) run <- read_rds(here::here("analysis", "output", "model_runs", "social_views.rds"))
if (charity_voluntarism == 1) run <- read_rds(here::here("analysis", "output", "model_runs", "charity_voluntarism.rds"))
if (demographics == 1) run <- read_rds(here::here("analysis", "output", "model_runs", "demographics.rds"))
if (public_political == 1) run <- read_rds(here::here("analysis", "output", "model_runs", "public_political.rds"))
if (public_political_social == 1) run <- read_rds(here::here("analysis", "output", "model_runs", "public_political_social.rds"))
if (public_political_social_charity == 1) run <- read_rds(here::here("analysis", "output", "model_runs", "public_political_social_charity.rds"))
if (public_political_social_charity_demo == 1) run <- read_rds(here::here("analysis", "output", "model_runs", "public_political_social_charity_demo.rds"))
data <- run$data
fit <- run$fit
```

# Model checking

```{r model-checking}
source(here::here("R", "stan_utility.R"))

# Gamma
draws_gamma <- fit %>% 
  extract(
    pars = str_c("Gamma[", 1:13, ",1]"),
    inc_warmup = TRUE,
    permuted = FALSE
  )

draws_gamma %>% 
  mcmc_hist_by_chain()

draws_gamma %>% 
  mcmc_trace(
    n_warmup = 2000,
    facet_args = list(
      nrow = 4, 
      labeller = label_parsed
    )
  )

# Extract the Omega parameters only and look at trace plots, etc.

# Omega
draws_Omega <- fit %>% 
  extract(
    regex_pars = c("Omega"),
    inc_warmup = TRUE,
    permuted = FALSE
  )

draws_gamma %>% 
  mcmc_hist_by_chain()

draws_gamma %>% 
  mcmc_trace(
    n_warmup = 2000,
    facet_args = list(
      nrow = 4, 
      labeller = label_parsed
    )
  )

# # Check diagnostics
# check_all_diagnostics(fit)

# Check for divergences
check_div(fit)

# # Check the effective sample size (NOT WORKING?)
# check_n_eff(fit)
# 
# # Check the Rhat statistic (NOT WORKING?)
# check_rhat(fit)

# Check trace plots (SUBSET?)
color_scheme_set("mix-blue-red")
fit %>%
  extract(
    inc_warmup = TRUE,
    permuted = FALSE
  ) %>%
  mcmc_trace(
    pars = str_c("Gamma[", 1:13, ",1]"),
    n_warmup = 2000,
    facet_args = list(
      nrow = 2, 
      labeller = label_parsed
    )
  )

# ggsave(
#   "intercept_trace_noncentered.png",
#   path = here::here("analysis", "output", "figures"),
#   width = 12, height = 6, units = "in"
# )
```

# Extract draws

```{r extract-draws}
# Extract posterior draws
draws <- fit %>% 
  spread_draws(Gamma[i, j]) %>% 
  ungroup() %>% 
  filter(.iteration > 500)

@@

# Plot marginals and contrasts

@@

intercept <- 0                            # Intercept-only
public_affairs <- 0                       # Public affairs
political_ideology <- 0                   # Political ideology
social_views <- 0                         # Social views
charity_voluntarism <- 1                  # Charity and voluntarism
demographics <- 0                         # Demographics
public_political <- 0                     # Public affairs + Political ideology
public_political_social <- 0              # + Social views
public_political_social_charity <- 0      # + Charity and voluntarism
public_political_social_charity_demo <- 0 # + Demographics

@@

>>>>

```{r load-packages, warning=FALSE, message=FALSE}
library(tidyverse)
library(donorheuristics)
library(rstan)
library(bayesplot)
library(tidybayes)
library(loo)
library(ggridges)
library(ggraph)
library(ggdag)
library(here)

@@

After selecting the model we'd like to run, model calibration proceeds as follows. Note that model calibration is run in a separate R script given how long it takes to run each model.

@@

The model is calibrated using a non-centered parameterization of a hierarchical multinomial logit model. Each model passed diagnostics checks, including the absence of divergent transitions, as part of model calibration.

@@

# Compute model fit and extract draws

The following analysis follows specifying the model we'd like to evaluate. Note that the output of the model calibration is too large to push to GitHub.

## Load model output

```{r load-data-output, warning=FALSE, message=FALSE, eval=FALSE}
# Load all model output
intercept_run <- read_rds(here::here("analysis", "output", "model_runs", "intercept_noncentered.rds"))
public_affairs_run <- read_rds(here::here("analysis", "output", "model_runs", "public_affairs.rds"))
political_ideology_run <- read_rds(here::here("analysis", "output", "model_runs", "political_ideology.rds"))
social_views_run <- read_rds(here::here("analysis", "output", "model_runs", "social_views.rds"))
charity_voluntarism_run <- read_rds(here::here("analysis", "output", "model_runs", "charity_voluntarism.rds"))
demographics_run <- read_rds(here::here("analysis", "output", "model_runs", "demographics.rds"))
public_political_run <- read_rds(here::here("analysis", "output", "model_runs", "public_political.rds"))
public_political_social_run <- read_rds(here::here("analysis", "output", "model_runs", "public_political_social.rds"))
public_political_social_charity_run <- read_rds(here::here("analysis", "output", "model_runs", "public_political_social_charity.rds"))
public_political_social_charity_demo_run <- read_rds(here::here("analysis", "output", "model_runs", "public_political_social_charity_demo.rds"))
```

## Compute model fit

```{r compute-fit, eval=FALSE}
# Compute fit using PSIS-LOO
intercept_loo <- loo(intercept_run$fit, save_psis = TRUE)
public_affairs_loo <- loo(public_affairs_run$fit, save_psis = TRUE)
political_ideology_loo <- loo(political_ideology_run$fit, save_psis = TRUE)
social_views_loo <- loo(social_views_run$fit, save_psis = TRUE)
charity_voluntarism_loo <- loo(charity_voluntarism_run$fit, save_psis = TRUE)
demographics_loo <- loo(demographics_run$fit, save_psis = TRUE)
public_political_loo <- loo(public_political_run$fit, save_psis = TRUE)
public_political_social_loo <- loo(public_political_social_run$fit, save_psis = TRUE)
public_political_social_charity_loo <- loo(public_political_social_charity_run$fit, save_psis = TRUE)
public_political_social_charity_demo_loo <- loo(public_political_social_charity_demo_run$fit, save_psis = TRUE)

# Compare model fit
loo_compare(intercept_loo, public_affairs_loo, political_ideology_loo, social_views_loo, charity_voluntarism_loo, demographics_loo, public_political_loo, public_political_social_loo, public_political_social_charity_loo, public_political_social_charity_demo_loo)
```

```
        elpd_diff se_diff
model1     0.0       0.0 
model3   -23.5       8.6 
model2   -37.5      12.9 
model6   -54.0      15.6 
model7   -70.2      15.4 
model4   -74.4      16.6 
model8  -163.9      23.6 
model5  -174.4      24.6 
model9  -401.3      38.4 
model10 -518.3      44.6
```

The intercept-only model works best in terms of PSIS-LOO, followed by models with political ideology, public affairs and demographics covariates. While PSIS-LOO doesn't properly account for the specific benefits of a hierarchical model, it suggests that political ideology above all other covariate sets specified, has the most explanatory power.

```{r save-fit, eval=FALSE}
# Create data frame
model_fit <- tibble(model = rep(NA, 10), elpd_loo = rep(NA, 10))

# Extract PSIS-LOO fit
model_fit$model[1] <- "intercept"; model_fit$loo[1] <- intercept_loo$elpd_loo
model_fit$model[2] <- "public_affairs"; model_fit$loo[2] <- public_affairs_loo$elpd_loo
model_fit$model[3] <- "political_ideology"; model_fit$loo[3] <- political_ideology_loo$elpd_loo
model_fit$model[4] <- "social_views"; model_fit$loo[4] <- social_views_loo$elpd_loo
model_fit$model[5] <- "charity_voluntarism"; model_fit$loo[5] <- charity_voluntarism_loo$elpd_loo
model_fit$model[6] <- "demographics"; model_fit$loo[6] <- demographics_loo$elpd_loo
model_fit$model[7] <- "public_political"; model_fit$loo[7] <- public_political_loo$elpd_loo
model_fit$model[8] <- "public_political_social"; model_fit$loo[8] <- public_political_social_loo$elpd_loo
model_fit$model[9] <- "public_political_social_charity"; model_fit$loo[9] <- public_political_social_charity_loo$elpd_loo
model_fit$model[10] <- "public_political_social_charity_demo"; model_fit$loo[10] <- public_political_social_charity_demo_loo$elpd_loo

# Save model output
write_rds(model_fit, here::here("analysis", "output", "model_fit.rds"))
```

## Load model fit and extract and save draws

```{r extract-draws, eval=FALSE}
intercept <- 1                            # Intercept-only
public_affairs <- 0                       # Public affairs
political_ideology <- 0                   # Political ideology
social_views <- 0                         # Social views
charity_voluntarism <- 0                  # Charity and voluntarism
demographics <- 0                         # Demographics
public_political <- 0                     # Public affairs + Political ideology
public_political_social <- 0              # + Social views
public_political_social_charity <- 0      # + Charity and voluntarism
public_political_social_charity_demo <- 0 # + Demographics

# Load and extract data and model output
if (intercept == 1) run <- read_rds(here::here("analysis", "output", "model_runs", "intercept_noncentered.rds"))
if (public_affairs == 1) run <- read_rds(here::here("analysis", "output", "model_runs", "public_affairs.rds"))
if (political_ideology == 1) run <- read_rds(here::here("analysis", "output", "model_runs", "political_ideology.rds"))
if (social_views == 1) run <- read_rds(here::here("analysis", "output", "model_runs", "social_views.rds"))
if (charity_voluntarism == 1) run <- read_rds(here::here("analysis", "output", "model_runs", "charity_voluntarism.rds"))
if (demographics == 1) run <- read_rds(here::here("analysis", "output", "model_runs", "demographics.rds"))
if (public_political == 1) run <- read_rds(here::here("analysis", "output", "model_runs", "public_political.rds"))
if (public_political_social == 1) run <- read_rds(here::here("analysis", "output", "model_runs", "public_political_social.rds"))
if (public_political_social_charity == 1) run <- read_rds(here::here("analysis", "output", "model_runs", "public_political_social_charity.rds"))
if (public_political_social_charity_demo == 1) run <- read_rds(here::here("analysis", "output", "model_runs", "public_political_social_charity_demo.rds"))
data <- run$data
fit <- run$fit

# Extract posterior draws
draws <- fit %>% 
  spread_draws(Gamma[i, j]) %>% 
  ungroup() %>% 
  filter(.iteration > 500)

@@

# Load draws and plot marginals and contrasts

@@

intercept <- 0                            # Intercept-only
public_affairs <- 0                       # Public affairs
political_ideology <- 0                   # Political ideology
social_views <- 0                         # Social views
charity_voluntarism <- 0                  # Charity and voluntarism
demographics <- 0                         # Demographics
public_political <- 0                     # Public affairs + Political ideology
public_political_social <- 0              # + Social views
public_political_social_charity <- 0      # + Charity and voluntarism
public_political_social_charity_demo <- 1 # + Demographics

@@

# # Create matrix of Gammas.
# gamma_draws <- draws %>% 
#   group_by(i, j) %>% 
#   summarize(gamma_mean = mean(Gamma))
# 
# gamma_matrix <- matrix(as.matrix(gamma_draws)[,3], nrow = max(draws$j), ncol = max(draws$i))
# write.csv(gamma_matrix, file = here::here("analysis", "output", "posterior_means.csv"))
